---
title: "Introduction to microbiomeDAAcomp"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to microbiomeDAAcomp}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The microbiomeDAAcomp package provides a comprehensive framework for comparing and evaluating different Differential Abundance Analysis (DAA) methods in microbiome studies. This vignette demonstrates the basic workflow and key functionalities of the package.

## Installation

You can install the development version of microbiomeDAAcomp from GitHub:

```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("cafferychen777/microbiomeDAAcomp")
```

## Basic Usage

### Data Import

First, let's import some microbiome data:

```{r eval=FALSE}
library(microbiomeDAAcomp)

# Import data (example with CSV format)
data <- import_data("path/to/data.csv", format = "csv")
```

### Data Simulation

For testing and evaluation purposes, we can simulate microbiome data:

```{r eval=FALSE}
# Simulate a dataset with 100 samples and 50 taxa
sim_data <- simulate_data(
  n_samples = 100,
  n_taxa = 50,
  dispersion = 0.3,
  effect_size = 1
)
```

### Running DAA Methods

The package allows you to run multiple DAA methods and compare their results:

```{r eval=FALSE}
# Run multiple DAA methods
results <- run_daa_methods(
  data = sim_data,
  methods = c("DESeq2", "ALDEx2")
)
```

## Future Development

This package is under active development. Upcoming features include:
- Implementation of additional DAA methods
- Advanced visualization tools
- Performance benchmarking capabilities
- Comprehensive method comparison metrics

## Session Info

```{r}
sessionInfo()
```
```