---
title: "Introduction to microbiomeDAAcomp"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to microbiomeDAAcomp}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

The microbiomeDAAcomp package provides a comprehensive framework for comparing and evaluating different Differential Abundance Analysis (DAA) methods in microbiome studies. This package addresses several key challenges in microbiome data analysis:

### Background

Microbiome studies often face challenges in identifying truly differentially abundant taxa between conditions due to:

- High dimensionality and sparsity of data
- Compositional nature of microbiome data
- Complex interactions between taxa
- Technical and biological variability

### Key Features

The package offers:

1. **Method Integration**: Implements multiple popular DAA methods including:

   - DESeq2 (Love et al., 2014)
   - ALDEx2 (Fernandes et al., 2014)
   - ANCOM-BC (Lin & Peddada, 2020)

2. **Performance Evaluation**:

   - Comprehensive metrics (sensitivity, specificity, precision)
   - Confidence interval estimation
   - Statistical power analysis
   - Method comparison frameworks

3. **Visualization Tools**:

   - Performance comparison plots
   - Method agreement visualization
   - Interactive plotting options
   - Customizable themes

4. **Statistical Analysis**:

   - Multiple testing correction
   - Effect size estimation
   - Robust statistical comparisons
   - Power analysis tools

## Setup

```{r setup}
# Load required packages
library(microbiomeDAAcomp)
```

## Basic Usage

### Create Sample Data

```{r}
# Create sample test results
set.seed(123)
n_samples <- 100
test_results <- list(
  "DESeq2" = sample(c(TRUE, FALSE), n_samples, replace = TRUE, prob = c(0.7, 0.3)),
  "edgeR" = sample(c(TRUE, FALSE), n_samples, replace = TRUE, prob = c(0.6, 0.4)),
  "limma" = sample(c(TRUE, FALSE), n_samples, replace = TRUE, prob = c(0.5, 0.5))
)
true_status <- sample(c(TRUE, FALSE), n_samples, replace = TRUE)
```

### Evaluate Performance

```{r}
# Calculate performance metrics
perf <- evaluate_performance(
  test_results,
  true_status,
  metrics = c("sensitivity", "specificity", "precision", "f1_score", "accuracy")
)

# Display results
print(perf)
```

### Visualize Results

```{r}
# Create different types of plots
# Heatmap visualization
plot_performance(perf, plot_type = "heatmap", theme = "default")

# Boxplot visualization
plot_performance(perf, plot_type = "boxplot", theme = "classic")

```

### Compare Methods

```{r}
# Create sample performance results
perf_data <- data.frame(
  method = rep(c("method1", "method2"), each = 10),
  accuracy = runif(20, 0.8, 0.95),
  precision = runif(20, 0.8, 0.95),
  recall = runif(20, 0.8, 0.95),
  f1_score = runif(20, 0.8, 0.95)
)

# Run comprehensive comparison
results <- compare_methods(perf_data, comparison_type = "comprehensive")

# Display results
print(results)
```

## Session Information

```{r}
sessionInfo()
```
